```{r setup, messages=F}
library(tidyverse)
library(ggplot2)
library(bbmle)

source(file = file.path(here::here(), "Scripts", "data_import.r"))
```

```{r}
dat
```




First look at available age composition data.

To me it looks like ages 1-4 are pretty underrepresented, most likely due to their low vulnerability to used gear.
Additionally, there's only one record of a 14yo fish, so we can exclude that. That leaves us with ages 5-13 for our age composition analysis
```{r}
dat %>% 
  ggplot(aes(age)) + 
  geom_histogram() + 
  theme_light() + 
  labs(title = "Age Distribution of Captured Arctic Grayling", x = "Age", y = "")
```



We'll also take a look at length at age.

Looks pretty standard, a good candidate for fitting a growth curve: no obvious outliers or artefacts
```{r}
dat %>% 
  ggplot(aes(age, length)) + 
  geom_point()
```

Let's try a first pass at fitting a growth curve
```{r}

# We'll need to create a subset of the data with no missing values
dat_vbl = 
  dat %>% 
  # We'll remove the first 4 years and the last 1 year of ages b/c of gear sensitivity and lack of data, respectively
  filter(age > 4, age < 14) %>% 
  select(age, length) %>% 
  drop_na()

# Define the functions we'll need for the optimizer

# We'll use the Von-Bertalanffey growth function
vbl = function(t, t_0, K, L_inf) {
  return(
    L_inf * (1 - exp(-K * (t - t_0)))
  )
}

# The negative log likelihood function for VBL
vbl_nll = function(obs_age, obs_len, sigma, t_0, K, L_inf, .log=TRUE) {
  pred_len = vbl(obs_age, t_0, K, L_inf)
  return(
    -1 * sum(
      dnorm(
        x = obs_len, 
        mean = pred_len,
        sd = sigma,
        log = .log
      ),
      na.rm = T
    )
  )
}
```
Let's get some estimates for our starting values
```{r}
# Define our guesses for the initial values
t_0_i = 0
K_i = .25
L_inf_i = 420 
sigma_i = 10

age_range = seq(min(dat$age, na.rm = T), max(dat$age, na.rm = T))
  
dat %>% 
  ggplot(aes(age, length)) + 
  geom_point() + 
  geom_line(
    data = data.frame(
      age = age_range, 
      length = vbl(age_range, t_0 = t_0_i, K = K_i, L_inf = L_inf_i)
    )
  )
```

That seems pretty decent, now let's run it through the optimizer
```{r}

# Now we can fit using a non-linear optimizer (a blazing fast guess ad checker)
fit = mle2(
  vbl_nll,
  data = list(
    obs_age = dat_vbl$age,
    obs_len = dat_vbl$length,
    log = T
  ),
  start = list(
    sigma = sigma_i,
    t_0 = t_0_i,
    K = K_i,
    L_inf = L_inf_i
  ),
  method="Nelder-Mead"
)

# Let's extract the estimated coefficients from our optimized fit
(sigma_p = coef(fit)['sigma'])
(t_0_p = coef(fit)['t_0'])
(K_p = coef(fit)['K'])
(L_inf_p = coef(fit)['L_inf'])

dat %>% 
  ggplot(aes(age, length)) + 
  geom_point() + 
  geom_line(
    data = data.frame(
      age = age_range, 
      length = vbl(age_range, t_0 = t_0_p, K = K_p, L_inf = L_inf_p)
    )
  ) + 
  geom_ribbon(
    data = data.frame(
      age = age_range, 
      length = vbl(age_range, t_0 = t_0_p, K = K_p, L_inf = L_inf_p)
    ),
    aes(
      ymin = length - 1.96 * sigma_p,
      ymax = length + 1.96 * sigma_p,
      fill = "95% C.I."
    ),
    alpha = 0.3
  ) +
  geom_ribbon(
    data = data.frame(
      age = age_range, 
      length = vbl(age_range, t_0 = t_0_p, K = K_p, L_inf = L_inf_p)
    ),
    aes(
      ymin = length - 0.674 * sigma_p,
      ymax = length + 0.674 * sigma_p,
      fill = "50% C.I."
    ),
    alpha = 0.3
  ) +
  # Write the estimates of the critical coefficients onto the plot
  annotate(
    geom="text", x=12, y=250, 
    label = paste(
      "Estimated Coefficients",
      # paste("Sigma:", sigma_p),
      # paste("t_0:", t_0_p),
      paste("Growth (K):", round(K_p, 4)),
      paste("Max Size (L_inf):", round(L_inf_p, 2)),
      sep = "\n"
    )
  ) +
  theme_light() + 
  labs(
    title = "Length-at-Age", 
    subtitle = "Von-Bertalanffey", 
    x = "Age (Years)", 
    y="Length (mm)"
  ) + 
  scale_fill_discrete(name = "Confidence Intervals")
```




